\section*{1.8.19}
(a) Daisy hears it?

\bigskip
\noindent
There are 64 possible outcomes.
Each outcome is a network with four nodes.
Let
\begin{eqnarray*}
A_1&=&\hbox{the set of all outcomes with arc $AD$} \\
A_2&=&\hbox{the set of all outcomes with arcs $AB$, $BD$, and not $AD$} \\
A_3&=&\hbox{the set of all outcomes with arcs $AC$, $CD$, and not $AD$} \\
A_4&=&\hbox{the set of one outcome with arcs $AB$, $BC$, and $CD$ only} \\
A_5&=&\hbox{the set of one outcome with arcs $AC$, $BC$, and $BD$ only}
\end{eqnarray*}
Each arc has probability $(1-p)$ hence
\begin{eqnarray*}
P(A_1)&=&1-p \\
P(A_2)&=&p(1-p)^2 \\
P(A_3)&=&p(1-p)^2 \\
P(A_4)&=&p^3(1-p)^3 \\
P(A_5)&=&p^3(1-p)^3
\end{eqnarray*}
All of the intersections are empty sets except
$$P(A_2\cap A_3)=p(1-p)^4$$
This intersection corresponds to the set of networks
with arcs $AB$, $AC$, $BD$, $CD$, and not $AD$.
Let $H$ be the event that $D$ hears it.
Then
\begin{eqnarray*}
P(H)&=&P(A_1\cup A_2\cup A_3\cup A_4\cup A_5) \\
&=&P(A_1)+P(A_2)+P(A_3)+P(A_4)+P(A_5)-P(A_2\cap A_3) \\
&=&1-p+2p(1-p)^2+2p^3(1-p)^3-p(1-p)^4
\end{eqnarray*}

\bigskip
\noindent
(b) Daisy hears it if Anne and Betty have quarreled?

\bigskip
\noindent
Let $B$ be the set of outcomes such that there is no arc $AB$.
Note that $P(B)=p$.
We have
\begin{eqnarray*}
P(B\cap A_1)&=&p(1-p) \\
P(B\cap A_2)&=&p^2(1-p)^2 \\
P(B\cap A_3)&=&0 \\
P(B\cap A_4)&=&0 \\
P(B\cap A_5)&=&p^3(1-p)^3
\end{eqnarray*}
Note that $B\cap A_5=A_5$ so $P(A_5)$ does not change.
Hence
\begin{eqnarray*}
P(H\mid B)&=&{P(
(B\cap A_1)\cup
(B\cap A_2)\cup
(B\cap A_5)
\over P(B)} \\
&=&1-p+p(1-p)^2+p^2(1-p)^3
\end{eqnarray*}

\bigskip
\noindent
(c) Daisy hears it if Betty and Chloe have quarreled?

\bigskip
\noindent
Let $B$ be the set of outcomes such that there is no arc $BC$.
Note that $P(B)=p$.
We have
\begin{eqnarray*}
P(B\cap A_1)&=&p(1-p) \\
P(B\cap A_2)&=&p^2(1-p)^2 \\
P(B\cap A_3)&=&p^2(1-p)^2 \\
P(B\cap A_4)&=&0 \\
P(B\cap A_5)&=&0
\end{eqnarray*}
In addtition
$$P((B\cap A_2)\cap(B\cap A_3))=p^2(1-p)^4$$
This intersection corresponds to the set of all networks with paths
$AB$, $AC$, $BD$, $CD$, and not $AD$ and not $BC$.
Hence
\begin{eqnarray*}
P(H\mid B)&=&{P(
(B\cap A_1)\cup
(B\cap A_2)\cup
(B\cap A_3)
\over P(B)} \\
&=&1-p+2p(1-p)^2-p(1-p)^4
\end{eqnarray*}

\bigskip
\noindent
(d) Daisy hears it if she has quarreled with Anne?

\bigskip
\noindent
Let $B$ be the set of outcomes such that there is no arc $AD$.
Note that $P(B)=p$.
We have
\begin{eqnarray*}
P(B\cap A_1)&=&0 \\
P(B\cap A_2)&=&p(1-p)^2 \\
P(B\cap A_3)&=&p(1-p)^2 \\
P(B\cap A_4)&=&p^3(1-p)^3 \\
P(B\cap A_5)&=&p^3(1-p)^3
\end{eqnarray*}
In addition
$$P((B\cap A_2)\cap(B\cap A_3))=p^2(1-p)^4$$
Hence
\begin{eqnarray*}
P(H\mid B)&=&{P(
(B\cap A_2)\cup
(B\cap A_3)\cup
(B\cap A_4)\cup
(B\cap A_5)
\over P(B)} \\
&=&2(1-p)^2+2p^2(1-p)^3-(1-p)^4
\end{eqnarray*}
