\section*{30.}
Luxco, a wholesale lightbulb manufacturer, has two factories. Factory A sells
bulbs in lots that consist of 1000 regular and 2000 softglow bulbs each. Random
sampling has shown that on the average there tend to be about 2 bad
regular bulbs and 11 bad softglow bulbs per lot. At factory B the lot size is
reversed---there are 2000 regular and 1000 softglow per lot---and there tend
to be 5 bad regular and 6 bad softglow bulbs per lot.

\bigskip
\noindent
The manager of factory A asserts, ``We're obviously the better producer; our
bad bulb rates are .2 percent and .55 percent compared to B's .25 percent and
.6 percent. We're better at both regular and softglow bulbs by half of a tenth
of a percent each.''

\bigskip
\noindent
``Au contraire,'' counters the manager of B, ``each of our 3000 bulb lots contains
only 11 bad bulbs, while A's 3000 bulb lots contain 13. So our .37
percent bad bulb rate beats their .43 percent.''

\bigskip
\noindent
Who is right?

\bigskip
\noindent
Let us use total probability to try and determine who is correct.
Let $R$ be the event that a regular bulb is manufactured and $S$ be
the event that a softglow bulb is manufactured.
Then factory A's probability of a bad bulb is
\begin{eqnarray*}
P(F)&=&P(F\mid R)P(R)+P(F\mid S)P(S)\\
&=&\left({2\over1000}\right)\left({1000\over3000}\right)
+\left({11\over2000}\right)\left({2000\over3000}\right)\\
&=&0.0043
\end{eqnarray*}
Factory B's probability of a bad bulb is
\begin{eqnarray*}
P(F)&=&P(F\mid R)P(R)+P(F\mid S)P(S)\\
&=&\left({5\over2000}\right)\left({2000\over3000}\right)
+\left({6\over1000}\right)\left({1000\over3000}\right)\\
&=&0.0037
\end{eqnarray*}
So it appears manager B is correct.
